---
author: "by Borja, Rachelle Erika"
title: "Mixed-Design ANOVA Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

This study examined whether dating preferences were influenced by:

- Gender (Male vs Female)
- Looks (Attractive, Average, Ugly)
- Personality (High Charisma, Some Charisma, No Charisma)

A mixed-design ANOVA was conducted to determine whether there was a three-way interaction among gender, looks, and personality on dating ratings.

# Load Required Libraries

```{r}
library(haven)
library(tidyverse)
library(afex)
library(emmeans)
library(rstatix)
library(ggplot2)
library(car)
```

# Import Dataset

```{r}
data <- read_sav("Looksorpersonality.sav")

head(data)
```

# Convert Data to Long Format

```{r}
long_data <- data %>%
  
  mutate(
    
    Subject = factor(1:n()),
    
    Gender = as.factor(Gender)
    
  ) %>%
  
  pivot_longer(
    
    cols = att_high:ug_none,
    
    names_to = "Condition",
    
    values_to = "Rating"
    
  )
```

# Create Looks and Personality Variables

```{r}
long_data <- long_data %>%
  
  mutate(
    
    Looks = case_when(
      str_detect(Condition, "att") ~ "Attractive",
      str_detect(Condition, "av") ~ "Average",
      str_detect(Condition, "ug") ~ "Ugly"
    ),
    
    Personality = case_when(
      str_detect(Condition, "high") ~ "High Charisma",
      str_detect(Condition, "some") ~ "Some Charisma",
      str_detect(Condition, "none") ~ "No Charisma"
    )
    
  )

long_data$Looks <- factor(
  long_data$Looks,
  levels = c("Attractive", "Average", "Ugly")
)

long_data$Personality <- factor(
  long_data$Personality,
  levels = c(
    "High Charisma",
    "Some Charisma",
    "No Charisma"
  )
)

long_data <- na.omit(long_data)

head(long_data)
```

# Descriptive Statistics

```{r}
descriptive_stats <- long_data %>%
  
  group_by(Gender, Looks, Personality) %>%
  
  summarise(
    
    Mean = round(mean(Rating), 3),
    
    SD = round(sd(Rating), 3),
    
    .groups = "drop"
    
  )

descriptive_stats
```

# Mixed-Design ANOVA

```{r}
anova_results <- aov_car(
  
  Rating ~ Gender * Looks * Personality +
    
    Error(Subject/(Looks * Personality)),
  
  data = long_data,
  
  factorize = FALSE
  
)

anova_results
```

# ANOVA Table

```{r}
nice(anova_results, es = "pes")
```

# Assumption Checks

## Normality of Residuals

```{r}
residuals_data <- residuals(anova_results$lm)

shapiro.test(residuals_data)
```

### Interpretation

- If p > .05, the residuals are approximately normally distributed.
- If p < .05, the residuals are not normally distributed.

# Homogeneity of Variance

```{r}
leveneTest(
  
  Rating ~ Gender * Looks * Personality,
  
  data = long_data
  
)
```

### Interpretation

- If p > .05, homogeneity of variance is satisfied.
- If p < .05, homogeneity of variance is violated.

# Estimated Marginal Means

```{r}
emmeans(
  
  anova_results,
  
  pairwise ~ Looks * Personality | Gender
  
)
```

# Interaction Plot

```{r}
means_data <- long_data %>%
  
  group_by(Gender, Looks, Personality) %>%
  
  summarise(
    
    Mean = mean(Rating),
    
    .groups = "drop"
    
  )

ggplot(
  
  means_data,
  
  aes(
    
    x = Looks,
    
    y = Mean,
    
    color = Personality,
    
    group = Personality
    
  )
  
) +

  geom_line(linewidth = 1) +
  
  geom_point(size = 3) +
  
  facet_wrap(~Gender) +
  
  labs(
    
    title = "Interaction Between Gender, Looks, and Personality",
    
    x = "Looks",
    
    y = "Mean Dating Rating"
    
  ) +
  
  theme_minimal(base_size = 14)
```

# Interpretation of Results

## Three-Way Interaction

The mixed-design ANOVA revealed a significant three-way interaction between gender, looks, and personality.

This indicates that the combined effects of attractiveness and charisma on dating ratings differed between males and females.

Specifically:

- Personality influenced dating ratings differently across attractiveness conditions.
- High charisma generally increased dating ratings.
- Personality appeared to compensate for lower physical attractiveness.
- The strength of these effects differed between genders.

# Significant Two-Way Interactions

## Gender × Looks

This interaction suggests that males and females evaluated physical attractiveness differently.

## Gender × Personality

This interaction indicates that charisma affected dating ratings differently depending on gender.

## Looks × Personality

This interaction suggests that personality influenced dating ratings differently across attractiveness levels.

# Main Effects

## Gender

The main effect of gender examines whether males and females differed overall in dating ratings.

## Looks

The main effect of looks examines whether attractiveness influenced dating ratings overall.

## Personality

The main effect of personality examines whether charisma influenced dating ratings overall.

# Overall Conclusion

The mixed-design ANOVA demonstrated that dating preferences were influenced by a combination of gender, physical attractiveness, and personality.

Overall, the personality had a strong influence on dating ratings.High charisma substantially improved dating evaluations. Personality compensated for lower physical attractiveness. The influence of attractiveness and personality differed between males and females.